Future Directions

### 2021-04-05 manuscripts/laplacian_smoothing/arXiv/minimax_optimal_laplacian_smoothing.tex
Responded to Siva's comments.

\begin{table}
	\begin{center}
		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
			\hline
			$d = 1$ & $\bm{n^{-2/3}}$ & \textcolor{red}{$\bm{n^{-2/3}}$} \\
			$d = 2,3$ & $\bm{n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
			$d = 4$ & $\bm{n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ &\textcolor{red}{$1$} \\
		\end{tabular}
	\end{center}
	\caption{Summary of estimation rates over first-order Sobolev balls. Black font marks new results from this paper, red font marks previously-known results; bold font marks minimax optimal rates. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal. Rates for thin-plate splines with $d \geq 2$ assume the estimator $\wt{f}$ interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}. We use ``$1$'' to indicate inconsistency (error not converging to 0). Furthermore, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
	\label{tbl:estimation_rates}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing_test} & Thin-plate splines \eqref{eqn:thin_plate_spline_test} \\
			\hline
			$d = 1$ & $\bm{n^{-4/5}}$ & \textcolor{red}{$\bm{n^{-4/5}}$} \\
			$d = 2,3$ & $\bm{n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
			$d \geq 4$ & $\bm{n^{-1/2}}$ & \textcolor{red}{$\bm{n^{-1/2}}$}
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}. Rates for thin-plate splines with $d \geq 2$ assume the test statistic $\wt{T}$ is computed with respect to an estimator $\wt{f}$ that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.  Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$. When $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
	\label{tbl:testing_rates}
\end{table}
%\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}

@@

To summarize: for $d = 1,2,$ or $3$, with high probability, the Laplacian smoothing estimator \smash{$\wh{f}$} has in-sample mean squared error that is within a constant factor of the minimax error. Some remarks:
\begin{itemize}
	\item The first-order Sobolev space $H^1(\Xset)$ does not continuously embed into $C^0(\Xset)$ when $d>1$ (in general, the $k$th order space $H^k(\Xset)$ does not continuously embed into $C^0(\Xset)$ except if $2k>d$). For this reason, one really cannot speak of pointwise evaluation of a Sobolev function $f_0 \in H^1(\Xset)$ when $d>1$ (as we do in Theorem~\ref{thm:laplacian_smoothing_estimation1} by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$). We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in Appendix~\ref{app:preliminaries}.
	% RJT: Alden, please add a section to the supplement to explain this part.
	% AJG: Added.
	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of at most a constant order).
	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency. The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but we note that as a practical matter one rarely chooses $r$ to be so large anyway.
	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm. When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart. In fact, as shown in Appendix~\ref{sec:graph_sobolev_seminorm}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}. We leave analysis of the Sobolev case to future work.
	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.
	%\sbcomment{Too mysterious a comment. Why is this sharper?} \\
	%\agcomment{Attempted to resolve.} 
\end{itemize}

@@

For a given $0 < \alpha < 1$, define a threshold \smash{$\wh{t}_{\alpha}$} as
\begin{equation*}
\wh{t}_{\alpha} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{1}{n}\sqrt{\frac{2}{\alpha} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
\end{equation*}
where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$L$}. The Laplacian smoothing test is then simply
\begin{equation*}
\wh{\varphi} = \1\bigl\{\wh{T} > \wh{t}_{\alpha}\bigr\}.
\end{equation*} 
%(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
We show in Appendix~\ref{sec:fixed_graph_error_bounds} that $\wh{f}$ is a level-$\alpha$ test. In the next theorem, we upper bound the worst-case risk $R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$ of $\wh{\varphi}$, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}. For this to hold, we will require a tighter range of scalings for the graph radius $r$.
\begin{enumerate}[label=(R\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:ls_kernel_radius_testing}
	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
	\begin{equation*}
	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{(d - 8)}{8 + 2d}} n^{\frac{d - 20}{32 + 8d}}.
	\end{equation*}
\end{enumerate}
We will also require that the radius of the Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
\begin{equation*}
M_{\max}(d) :=
\begin{cases*}
n^{1/8} & \textrm{$d = 1$} \\
n^{(4 - d)/(4d)} & \textrm{$d \geq 2$}.
\end{cases*}
\end{equation*}

We now give Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
\begin{theorem}
	\label{thm:laplacian_smoothing_testing}
	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and where $M \leq M_{\max}(d)$. Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}. There are constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing}
	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
	\end{equation} 
	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$.
\end{theorem}
%\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity). Otherwise you've introduced that terminology for no reason.}

Some remarks:
\begin{itemize}
	\item As mentioned earlier, Sobolev balls $H^1(\Xset,M)$ for $d \geq 4$ include quite irregular functions $f \not\in \Leb^4(\Xset)$. Proving tight lower bounds in this case is nontrivial, and as far as we understand such an analysis remains outstanding. On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$. Moreover, by setting $\rho = 0$ so that the resulting estimator $\wh{f}$ interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound. That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C(b^2 + \sqrt{1/\alpha})n^{-1/2}$}, we have that \smash{$\Ebb_0[\wh{\varphi}] \leq \alpha$} and
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing_low_smoothness}
	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2}.
	\end{equation} 
	\item To compute the data-dependent threshold \smash{$\wh{t}_{\alpha}$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$. Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time). But in practice we would not recommend using \smash{$\wh{t}_{\alpha}$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}. Recent work \cite{kim2020minimax}, has shown that in a variety of closely related settings, calibration of a test statistic via the permutation test often retains minimax-optimal power, and we expect similar results to hold for the Laplacian smoothing-based test statistic.

@@

With some results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent. First, we elaborate on why this difference in performance is so surprising. As mentioned previously, the penalties in \eqref{eqn:laplacian_smoothing}, \eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for $f \in C^2(\Xset)$, 
\begin{equation}
\label{eqn:seminorm_consistency}
\begin{aligned}
\lim \frac{1}{n^2 r^{d + 2}} f^\top L f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
& = \int_{\Xset} \|\nabla f(x)\|_2^2 p^2(x) \,dx.
\end{aligned}
\end{equation}
In the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
\begin{equation*}
\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f),
\end{equation*}
and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies e.g., Dirichlet boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator $\wh{f}$ and the thin-plate spline estimator $\wt{f}$ are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$). But it does seem to suggest that the two estimators should behave somewhat similarly.

@@

\begin{theorem}
	\label{thm:laplacian_smoothing_estimation_manifold}
	As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$. There are constants $N,C,c>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
	\begin{equation*}
	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)},
	\end{equation*}
	with probability at least $1 - \delta -  C n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
	% RJT: Alden, please verify that what I wrote is correct.
	% AJG: Verified.
\end{theorem}
In a similar vein, we obtain results for manifold adaptive testing under the following condition on the graph radius parameter:
\begin{enumerate}[label=(R\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp:ls_kernel_radius_testing_manifold}
	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
	\begin{equation*}
	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge M^{\frac{(m - 8)}{8 + 2m}} n^{\frac{m - 20}{32 + 8m}}.
	\end{equation*}
\end{enumerate}

\begin{theorem}
	\label{thm:laplacian_smoothing_testing_manifold}
	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing_manifold}
	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\Bigl(b^2 + b\sqrt{\frac{1}{\alpha}}\Bigr) ,
	\end{equation} 
	then the worst-case risk is upper bounded, $R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$.
	% RJT: Alden, please verify that what I wrote is correct.
	% AJG: Verified.
\end{theorem}
% \sbcomment{Same comment as before on what to emphasize.}

>>>>

\begin{table}
	\begin{center}
		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing} & Thin-plate splines \eqref{eqn:thin_plate_spline} \\
			\hline
			$d = 1$ & $\bm{n^{-2/3}}$ & \textcolor{red}{$\bm{n^{-2/3}}$} \\
			$d = 2,3$ & $\bm{n^{-2/(2 + d)}}$ & \textcolor{red}{$1$} \\
			$d = 4$ & $\bm{n^{-1/3}} (\log n)^{1/3}$ & \textcolor{red}{$1$} \\
			$d \geq 5$  & $(\log n/n)^{4/(3d)}$ &\textcolor{red}{$1$} \\
		\end{tabular}
	\end{center}
	\caption{Summary of estimation rates over first-order Sobolev balls. Black font marks new results from this paper, red font marks previously-known results; bold font marks minimax optimal rates. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius of the Sobolev ball is also optimal. The rates for thin-plate splines with $d \geq 2$ assume the estimator \smash{$\wt{f}$} interpolates the responses, \smash{$\wt{f}(X_i) = Y_i$} for $i = 1,\ldots,n$; see the discussion in Section~\ref{sec:summary}. Here, we use ``$1$'' to indicate inconsistency (error not converging to 0). Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all Laplacian smoothing results hold with $d$ replaced by $m$, without any change to the method itself.}
	\label{tbl:estimation_rates}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{p{.1\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Dimension & Laplacian smoothing \eqref{eqn:laplacian_smoothing_test} & Thin-plate splines \eqref{eqn:thin_plate_spline_test} \\
			\hline
			$d = 1$ & $\bm{n^{-4/5}}$ & \textcolor{red}{$\bm{n^{-4/5}}$} \\
			$d = 2,3$ & $\bm{n^{-4/(4 + d)}}$ & \textcolor{red}{$n^{-1/2}$} \\
			$d \geq 4$ & $\bm{n^{-1/2}}$ & \textcolor{red}{$\bm{n^{-1/2}}$}
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over first-order Sobolev balls; black, red, and bold fonts are used as in Table~\ref{tbl:estimation_rates}. The rates for thin-plate splines with $d \geq 2$ assume the test statistic \smash{$\wt{T}$} is computed using an \smash{$\wt{f}$} that interpolates the responses, $\wt{f}(X_i) = Y_i$ for $i = 1,\ldots,n$.  Rates for $d \geq 4$ assume that $f_0 \in \Leb^4(\Xset,M)$. Lastly, when $\Xset$ is an $m$-dimensional manifold embedded in $\Rd$, all rates hold with $d$ replaced by $m$.}
	\label{tbl:testing_rates}
\end{table}
%\footnotetext{Assuming $f_0 \in \Leb^4(\Xset,M)$.}

@@

To summarize: for $d = 1,2,$ or $3$, with high probability, the Laplacian smoothing estimator \smash{$\wh{f}$} has in-sample mean squared error that is within a constant factor of the minimax error. Some remarks:
\begin{itemize}
	\item The first-order Sobolev space $H^1(\Xset)$ does not continuously embed into $C^0(\Xset)$ when $d>1$ (in general, the $k$th order space $H^k(\Xset)$ does not continuously embed into $C^0(\Xset)$ except if $2k>d$). For this reason, one really cannot speak of pointwise evaluation of a Sobolev function $f_0 \in H^1(\Xset)$ when $d>1$ (as we do in Theorem~\ref{thm:laplacian_smoothing_estimation1} by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$). We can resolve this by appealing to what are known as \emph{Lebesgue points}, as explained in Appendix~\ref{app:preliminaries}.
	% RJT: Alden, please add a section to the supplement to explain this part.
	% AJG: Added.
	\item The assumption $M \leq n^{1/d}$ ensures that the upper bound provided in the theorem is meaningful (i.e., ensures it is of at most a constant order).
	\item The lower bound on $r$ imposed in condition \ref{asmp:ls_kernel_radius_estimation} is compatible with practice, where by far the most common choice of radius is the connectivity threshold \smash{$r \asymp (\log(n)/n)^{1/d}$}, which makes \smash{$G_{n,r}$} as sparse as possible while still being connected, for maximum computational efficiency. The upper bound may seem a bit more mysterious---we need it for technical reasons to ensure that \smash{$\wh{f}$} does not overfit, but we note that as a practical matter one rarely chooses $r$ to be so large anyway.
	\item It is possible to extend \smash{$\wh{f}$} to be defined on all of $\Xset$ and then evaluate the error of such an extension (as measured against $f_0$) in $\Leb^2(\Xset)$ norm. When \smash{$\wh{f}$} and $f_0$ are suitably smooth, tools from empirical process theory (see e.g., Chapter 14 of \citet{wainwright2019}) or approximation theory (e.g., Section 15.5 of \citet{johnstone2011}) guarantee that the $\Leb^2(\Xset)$ error is not too much greater than its in-sample counterpart. In fact, as shown in Appendix~\ref{subsec:laplacian_smoothing_estimation1_pf}, if $f_0$ is Lipschitz smooth and we extend \smash{$\wh{f}$} to be piecewise constant over the Voronoi tessellation induced by $X_1,\ldots,X_n$, then the out-of-sample error \smash{$\|\wh{f}-f_0\|_{\Leb^2(\Xset)}$} is within a negligible factor of the in-sample error \smash{$\|\wh{f}-f_0\|_n$}. We leave analysis of the Sobolev case to future work.
	\item When $f_0$ is Lipschitz smooth, we can also replace the factor of $\delta$ in the high probability bound by a factor of $\delta^2/n$, which is always smaller than $\delta$ when $\delta \in (0,1)$.
	%\sbcomment{Too mysterious a comment. Why is this sharper?} \\
	%\agcomment{Attempted to resolve.} 
\end{itemize}

@@

For a given $0 < \alpha < 1$, define a threshold \smash{$\wh{t}_{\alpha}$} as
\begin{equation*}
\wh{t}_{\alpha} = \frac{1}{n}\sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^2} + \frac{1}{n}\sqrt{\frac{2}{\alpha} \sum_{k = 1}^{n} \frac{1}{(\rho \lambda_k + 1)^4}},
\end{equation*}
where we recall $\lambda_k$ is the $k$th smallest eigenvalue of \smash{$L$}. The Laplacian smoothing test is then simply
\begin{equation*}
\wh{\varphi} = \1\bigl\{\wh{T} > \wh{t}_{\alpha}\bigr\}.
\end{equation*} 
%(Clearly \smash{$\wh{\varphi}$} depends on $b$, but we suppress this notationally.) 
We show in Appendix~\ref{sec:fixed_graph_error_bounds} that \smash{$\wh{f}$} is a level-$\alpha$ test. In the next theorem, we upper bound the worst-case risk \smash{$R_n(\wh{\varphi},H^1(\mc{X},M),\epsilon)$} of \smash{$\wh{\varphi}$}, whenever $\epsilon$ is at least (a constant times) the critical radius given in~\eqref{eqn:sobolev_space_testing_critical_radius}. For this to hold, we will require a tighter range of scalings for the graph radius $r$.
\begin{enumerate}[label=(R\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:ls_kernel_radius_testing}
	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies
	\begin{equation*}
	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{d}} \leq r \leq c_0 \wedge M^{\frac{(d - 8)}{8 + 2d}} n^{\frac{d - 20}{32 + 8d}}.
	\end{equation*}
\end{enumerate}
We will also require that the radius of the Sobolev class not be too large. Precisely, we will require $M \leq M_{\max}(d)$, where we define
\begin{equation*}
M_{\max}(d) :=
\begin{cases*}
n^{1/8} & \textrm{$d = 1$} \\
n^{(4 - d)/(4d)} & \textrm{$d \geq 2$}.
\end{cases*}
\end{equation*}

We now give Theorem~\ref{thm:laplacian_smoothing_testing}, our main testing result.
\begin{theorem}
	\label{thm:laplacian_smoothing_testing}
	Given i.i.d.\ draws $(X_i,Y_i)$, $i=1,\ldots,n$ from \eqref{eqn:signal_plus_noise_model}, assume $f_0 \in H^1(\Xset,M)$ where $\Xset \subseteq \Rd$ with $d < 4$, and $M \leq M_{\max}(d)$. Assume \ref{asmp:domain}, \ref{asmp:density} on the design distribution $P$, and assume \smash{$G_{n,r}$} is computed with a kernel $K$ satisfying \ref{asmp:kernel}. There exist constants $N,C,C_1,c_1>0$ such that for any $n \geq N$, and any radius $r$ as in \ref{asmp:ls_kernel_radius_testing}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{d + 2})^{-1} n^{-4/(4 + d)} M^{-8/(4 + d)}$}, satisfies the following: for any $b \geq 1$, if
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing}
	\epsilon^2 \geq C M^{2d/(4 + d)} n^{-4/(4 + d)}\biggl(b^2 + b\sqrt{\frac{1}{\alpha}}\biggr) ,
	\end{equation} 
	then the worst-case risk satisfies the upper bound: \smash{$R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + C_1n\exp(-c_1nr^d)\bigr)$}.
\end{theorem}
%\sbcomment{Personally, I wouldn't emphasize the Type II error so much in the Theorem (it is not tight and it's not an interesting quantity anyway), and write \eqref{eqn:laplacian_smoothing_testing} more directly as an upper bound on the critical radius (which is the interesting quantity). Otherwise you've introduced that terminology for no reason.}

Some remarks:
\begin{itemize}
	\item As mentioned earlier, Sobolev balls $H^1(\Xset,M)$ for $d \geq 4$ include quite irregular functions $f \not\in \Leb^4(\Xset)$. Proving tight lower bounds in this case is nontrivial, and as far as we understand such an analysis remains outstanding. On the other hand, if we explicitly assume that $f_0 \in \Leb^4(\Xset,M)$, then \citet{guerre02} show that the testing problem is characterized by a dimension-free lower bound $\epsilon^{2}(\Leb^4(\Xset,M)) \gtrsim n^{-1/2}$. Moreover, by setting $\rho = 0$ so that the resulting estimator \smash{$\wh{f}$} interpolates the responses $Y_1,\ldots,Y_n$, the subsequent test \smash{$\wh{\varphi}$} will achieve (up to constants) this lower bound. That is, for any $f_0 \in \Leb^4(\Xset,M)$ such that \smash{$\|f_0\|_{\Leb^2(\Xset)}^2 \geq C(b^2 + \sqrt{1/\alpha})n^{-1/2}$}, we have that \smash{$\Ebb_0[\wh{\varphi}] \leq \alpha$} and
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing_low_smoothness}
	\Ebb_{f_0}\bigl[1 - \wh{\varphi}\bigr] \leq \frac{C(1 + M^4)}{b^2}.
	\end{equation} 
	\item To compute the data-dependent threshold \smash{$\wh{t}_{\alpha}$}, one must know all of the eigenvalues $\lambda_1,\ldots,\lambda_n$. Computing all these eigenvalues is far more expensive (cubic-time) than computing \smash{$\wh{T}$} in the first place (nearly-linear-time). But in practice we would not recommend using \smash{$\wh{t}_{\alpha}$} anyway, and would instead we make the standard recommendation to calibrate via a permutation test \citep{hoeffding1952}. Recent work \cite{kim2020minimax}, has shown that in a variety of closely related settings, calibration of a test statistic via the permutation test often retains minimax-optimal power, and we expect similar results to hold for the Laplacian smoothing-based test statistic.

@@

With some results in hand, let us pause to offer some explanation of why Laplacian smoothing can be optimal in settings where thin-plate splines are not even consistent. First, we elaborate on why this difference in performance is so surprising. As mentioned previously, the penalties in \eqref{eqn:laplacian_smoothing}, \eqref{eqn:thin_plate_spline} can be closely tied together: \citet{bousquet03} show that for $f \in C^2(\Xset)$, 
\begin{equation}
\label{eqn:seminorm_consistency}
\begin{aligned}
\lim \frac{1}{n^2 r^{d + 2}} f^\top L f & = \int_{\Xset} f(x) \cdot \Delta_Pf(x) p(x) \,dx \\
& = \int_{\Xset} \|\nabla f(x)\|_2^2 p^2(x) \,dx.
\end{aligned}
\end{equation}
In the above, the limit is as $n \to \infty$ and $r \to 0$, $\Delta_P$ is the (weighted) Laplace-Beltrami operator
\begin{equation*}
\Delta_Pf := -\frac{1}{p} \dive\bigl(p^2\nabla f),
\end{equation*}
and the second equality follows using integration by parts.\footnote{Assuming $f$ satisfies Neumann boundary conditions.} To be clear, this argument does not formally imply that the Laplacian eigenmaps estimator \smash{$\wh{f}$} and the thin-plate spline estimator \smash{$\wt{f}$} are close (for one, note that \eqref{eqn:seminorm_consistency} holds for $f \in C^2(\Xset)$, whereas the optimization in \eqref{eqn:thin_plate_spline} considers a much broader set of continuous functions with weak derivatives in $\Leb^2(\Xset)$). But it does seem to suggest that the two estimators should behave somewhat similarly.

@@

\begin{theorem}
	\label{thm:laplacian_smoothing_estimation_manifold}
	As in Theorem \ref{thm:laplacian_smoothing_estimation1}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}, and $M \leq n^{1/m}$. There are constants $N,C,c>0$ (not depending on $f_0$) such that for any \smash{$n \geq N$}, and any $r$ as in \ref{asmp:ls_kernel_radius_estimation_manifold}, the Laplacian smoothing estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$L = L_{n,r}$} and \smash{$\rho = M^{-4/(2 + m)} (nr^{m + 2})^{-1} n^{-2/(2 + m)}$}, satisfies
	\begin{equation*}
	\bigl\|\wh{f} - f_0\bigr\|_n^2 \leq \frac{C}{\delta} M^{2m/(2 + m)} n^{-2/(2 + m)},
	\end{equation*}
	with probability at least $1 - \delta -  C n\exp(-cnr^m) - \exp(-c(M^2n)^{m/(2+m)})$.
	% RJT: Alden, please verify that what I wrote is correct.
	% AJG: Verified.
\end{theorem}
In a similar vein, we obtain results for manifold adaptive testing under the following condition on the graph radius parameter.
\begin{enumerate}[label=(R\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp:ls_kernel_radius_testing_manifold}
	For constants $C_0,c_0>0$, the neighborhood graph radius $r$ satisfies 
	\begin{equation*}
	C_0\biggl(\frac{\log n}{n}\biggr)^{\frac{1}{m}} \leq r \leq c_0 \wedge M^{\frac{(m - 8)}{8 + 2m}} n^{\frac{m - 20}{32 + 8m}}.
	\end{equation*}
\end{enumerate}

\begin{theorem}
	\label{thm:laplacian_smoothing_testing_manifold}
	As in Theorem \ref{thm:laplacian_smoothing_testing}, but where $\Xset \subseteq \Rd$ is a manifold with intrinsic dimension $m < 4$, $M \leq M_{\max}(m)$, and the design distribution $P$ obeys \ref{asmp:domain_manifold}, \ref{asmp:density_manifold}.
	There are constants $N,C,c>0$ such that for any $n \geq N$, and any $r$ as in \ref{asmp:ls_kernel_radius_testing_manifold}, the Laplacian smoothing test \smash{$\wh{\varphi}$} based on the estimator \smash{$\wh{f}$} in \eqref{eqn:laplacian_smoothing}, with \smash{$\rho = (nr^{m + 2})^{-1} n^{-4/(4 + m)} M^{-8/(4 + m)}$}, satisfies the following: for any $b \geq 1$, if
	\begin{equation}
	\label{eqn:laplacian_smoothing_testing_manifold}
	\epsilon^2 \geq C M^{2m/(4 + m)} n^{-4/(4 + m)}\biggl(b^2 + b\sqrt{\frac{1}{\alpha}}\biggr) ,
	\end{equation} 
	then the worst-case risk satisfies the upper bound: \smash{$R_n(\wh{\varphi},H^1(\mc{X},M), \epsilon) \leq C/b + Cn\exp(-cnr^m)$}.
	% RJT: Alden, please verify that what I wrote is correct.
	% AJG: Verified.
\end{theorem}
% \sbcomment{Same comment as before on what to emphasize.}
